% !TeX program = xelatex
% !TeX encoding = UTF-8
% !TeX spellcheck = it_IT

\begin{filecontents}{\jobname.bib}
@manuscript{Vatgr1,
  collection     = {Vaticanus graecus},
  location       = {Città del Vaticano},
  library        = {Bibliotheca Apostolica Vaticana},
  shelfmark      = {1},
  bookpagination = {folio},
  columns        = {1},
  dating         = {X saec.},
  shortlibrary   = {Vat.},
  pagetotal      = {300},
  shorthand      = {O},
  support        = {pergament},
  keywords       = {manuscripti},
}

@manuscript{Pargr1807,
  collection     = {Parisinus Graecus},
  location       = {Paris},
  library        = {Bibliothèque Nationale de France},
  shelfmark      = {1807},
  bookpagination = {folio},
  columns        = {2},
  dating         = {IX saec.},
  shortlibrary   = {Par.},
  pagetotal      = {300},
  shorthand      = {A},
  support        = {pergament},
  keywords       = {manuscripti},
}

@manuscript{Beinecke278,
  collection = {Beinecke},
  location   = {New Haven (Connecticut)},
  library    = {Yale University},
  shelfmark  = {278},
  columns    = {2},
  dating     = {4 October 1453},
  pages      = {1\recto--160\recto},
  pagetotal  = {164},
  pagination = {column},
  support    = {paper},
  annotation = {Written in Italy by Ioannes Skoutariotes in 1453 (see Vogel and Gardthausen, pp. 197-99; Canart, p. 68; de Meyier, p. 262). Acquired from Frater Philippus Lapaccinus by the Dominican convent of San Marco in Florence in the 15th century; ownership inscription of 15th century on f. ii verso is partially erased but still legible under ultraviolet light: "Iste liber est conuentus Sancti Marci de Florentia ordinis predicatorum. In bancho vj occidentis librarie grece. Habitus a fratre philippo lapaccino filio natiuo. N. LXXXVIIII." (B. Ullman and P. Stadter, The Public Library of Renaissance Florence [Padua, 1972] pp. 76, n. 1; 264; 279; 317.) Collection of Sir Thomas Phillipps (no. 10371, on spine), who purchased it from Payne (Cat. 1835, no. 100). Acquired in 1953 by Thomas E. Marston (bookplate) from C. A. Stonehill. Gift of Thomas E. Marston in 1959.},
  shorthand      = {B},
  keywords   = {manuscripti},
  origin     = {Italy},
  scribe     = {Ioannes Skoutariotes},
}
\end{filecontents}

\documentclass{article}
\usepackage{filecontents}

\usepackage{polyglossia}
	\setmainlanguage{french}
\usepackage{libertinus-otf}

\usepackage[
	backend=biber,%
	language=italian,%
	autolang=other,%
	bibstyle=manuscripts-noautoshorthand,
	]{biblatex}
\addbibresource{\jobname.bib}

\begin{document}
\cite{Vatgr1}

%\shcite{Vatgr1}

\nocite{*}
\printshorthands[type=manuscript,title=Conspectus siglorum]

\printshorthands[type=manuscript,env=details,title=Description of manuscripts]
\end{document}
